---
title: "Mental Health Admissions by Moon Cycle"
author: "Wellsin Price"
date: "2023-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(88)
```

# Mental Health Admissions and Full Moons

## Data analysis reproducing findings of Blackman, S., and Catalina, D.(1973).

The purpose of this script is to reproduce Blackman and Catalina's conclusion by way of computational analysis. The motivation is to demonstrate both programming and statistical thinking skills.

### Load Libraries and dataset

```{r}
library(tidyverse) # Data Science toolbox
library(Stat2Data) # Dataset
library(infer) # For statistical analysis
```

```{r}
data("MentalHealth") # Loaded from Stat2Data
MentalHealth # Observe the imported dataset
```

### Data Exploration and Wrangling
The relationships between admissions and the variables month and moon need to be explored to determine which inquiry is worth exploring further.
The script below wrangles the data to do such that.
```{r}
# Piping the dataset to compare the mean admissions to the moon cycle
MentalHealth %>% 
  group_by(Moon) %>%
  summarize(admission_mean = mean(Admission))
# Piping the dataset to compare mean admissions to month
MentalHealth %>%
  group_by(Month) %>%
  summarize(admission_mean = mean(Admission))
```

From this, we can establish that the relationship between Moon and Admission provide a more straightforward analysis given less entries to compare. Observing the data trends between the two new tables also alludes to two different stories to be told.

### Analysis into Month:
April demonstrated the highest admission mean of the 12 months, followed by May and March respectively. This could demonstrate an interesting phenomena: post-winter blues. 


### Analysis into Moon:
The study aimed to understand the relationship between admissions and moon cycles, so naturally an analysis replicating theirs is in order.
```{r}
# Visualizing the dataset and two variables as a box plot
ggplot(MentalHealth, aes(x = Moon, y = Admission, fill = Moon)) +
  geom_boxplot() +
  labs(title = "Boxplot of Admissions to Mental Health ERs During Moon Cycle")
```

X

```{r}
# Filtering out After and Before separately to compare to During
# After:
mentalhealth_after_during <- MentalHealth %>%
  filter(Moon != "Before")
# Before:
mentalhealth_before_during <- MentalHealth %>%
  filter(Moon != "After")

```

```{r}
# Calculation of observed difference in means
# After:
obs_diff_after_during <- mentalhealth_after_during %>%
  specify(Admission ~ Moon) %>%
  calculate(stat = "diff in means", order = c("After", "During"))
# Before:
obs_diff_before_during <- mentalhealth_before_during %>%
  specify(Admission ~ Moon) %>%
  calculate(stat = "diff in means", order = c("Before", "During"))
# View calculated values
obs_diff_after_during
obs_diff_before_during
```

Given the smaller size, after-during is better suited for hypothesis testing

```{r}
# Generate null distribution of delta
adm_in_null <- mentalhealth_after_during %>%
  specify(Admission ~ Moon) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 5000, type = 'permute')
```

```{r}
# Calculate difference in means
adm_diff_in_null <- adm_in_null %>%
  calculate(stat = "diff in means", order = c("After", "During"))
# Slice a window of 5 replicates
adm_diff_in_null %>%
  slice(1:5)
```

```{r}
# Get the p-value for the calculated difference
adm_diff_in_null %>%
  get_pvalue(obs_stat = obs_diff_after_during, direction = "both")
```

```{r}
# Run a t test for comparison
t.test(Admission ~ Moon, var.equal = TRUE, data = mentalhealth_after_during)
```

```{r}
# Fit data to regression model
mentalhealth_regression <- lm(formula = Admission ~ Moon, data = MentalHealth)
# Get regression table
# get_regression_table(mentalhealth_regression)
```
